Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex

### 2019-03-27 Manuscript.Rmd
advances on methods revision

For all our analyses, we defined our search space based on the 17 network parcellation proposed by Yeo et al. [-@Yeo2011]. First, We selected vertices on the cortical surface that were contained by the DMN and limbic networks along the medial wall. Next, we overlayed these networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained parcels that shared a considerable amount of vertices with them (visually inspected). This resulted in 45 regions of interest (ROI) across hemispheres (Table 1). These ROIs were naturally divided into two spatially discontiguous clusters that covered PCC and mPFC, thus facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels were differentially associated with valuation or DMN effects. For subjective value, we gathered peak brain coordinates of activity from 198 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DMN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the surviving areas post-statistical thresholding from each study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fs_average, 32,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl2012] (http://surfer.nmr.mgh.harvard.edu/). Any sphere sections that overlapped with the cortex were mapped to our ROIs, and we tallied how many studies had positive indicator values in each parcel.

With these lists, we identified areas that were significantly more associated to either literature. We permuted the label of every study (i.e. DMN or valuation) 5000 times while preseving the total number of studies observed in each domain, and on each iteration stored the maximum resulting statistic (across all parcels) from chi-squared tests of proportions that compared the number of studies that contained each area. This gave us a null distribution of 5000 chi-squared values. The 95th percentile of this distribution served as an FWE-corrected significance threshold to evaluate unpermuted chi-squared values.

@@

Our fMRI analyses used resting-state fMRI data from the Human Connectome Project [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state runs (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) each comprised 1200 time points(14 min 33 s) for a total of 4800 time points. Two runs were acquired on each day, with the phase encoding direction set to left-right for one run and right-left for the other. Only subjects with both left-right and right-left phase encoding were included (i.e. subjects with four rsfMRI sessions). In addition, only datasets with either low motion levels (under 1.5mm) or under 0.5 mm mean framewise displacement (FD) were used.

Data initially underwent the HCP minimally preprocessing pipeline [@Glasser2013], which included considerable gradient nonlinearity correction, motion correction, EPI distortion correction, temporal denoising based on independent components (FIX, [@Griffanti2014; @Salimi-Khorshidi2014]), high-pass filtering (0.0005 Hz threshold), and MNI152-based normalization. Data were further preprocessed using an in-house pipeline described previously [@Tobyne2017]. Steps included linear interpolation across timepoints with over 0.5 mm of framewise displacement, band-pass filtering (allowed frequencies ranged from 0.009 and 0.08 Hz), temporal denoising and high-motion censoring by deletion, as well as mean greyordinate signal regression [@Burgess2016]. Volumes that displayed an FD of over 0.5 mm were considered as spikes and removed from the subject's dataset, and subjects whose spike counts were above 5% of their total data were excluded from analyses. Finally, data acquired on the same day (i.e. left-right and right-left phase encoding session) were temporally demeaned. Each subject's brain was comprised of 32k standard HCP greyordinates per hemisphere (combined in a CIFTI file). We retained only the cortical surfaces, which resulted in 59,412 total surface vertices per subject.

### *Within-ROI Network Definition*

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the ROIs (n = 4,801 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. We report the proportion of retained edges (i.e. network density) in the Results section. As part of the evaluation step, we produced these network weight matrices (and the community detection outlined below) at four levels: 1) for each run separately (1200 TRs each); 2) for the concatenated time series for the two runs on the same day (2400 TRs total); 3) for each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

We performed a coordinate-based meta-analysis to identify unique standard regions across DMN and SV literatures. Volumetric coordinates from a set of DMN and SV studies were projected onto a cortical surface, and mapped to a multimodal cortical parcellation [@Glasser2016] to produce a list of brain areas reported per study. The 45 areas considered were limited to the medial portion of the default mode and limbic networks, as defined by the Yeo et al.'s [-@Yeo2011] 17 network parcellation. Unique areas were identified by first permuting the affiliation of each study (i.e. DMN or SV) to create a distribution of chi-squared statistics based on the comparison of the number of reports per area between literatures (see Methods for details). This putative null distribution was used to identify regions more often reported in either literature.

>>>>

For all analyses, we defined our search space based on the 17-network parcellation proposed by Yeo et al. [-@Yeo2011]. First, we selected vertices on the cortical surface that were contained by the DMN and limbic networks along the medial wall. Next, we overlayed those networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained a set of parcels that covered approximately the same brain regions. This resulted in a search space that consisted of 40 parcels across hemispheres (Table 1). The search space was naturally divided into two spatially discontiguous clusters that covered PCC and mPFC, facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels within the search space were preferentially associated with valuation or DMN effects. For subjective valuation, we gathered peak brain coordinates of activity from 200 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DMN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the areas that exceeded the statistical significance threshold in each original study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fs_average, 32,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl2012] (http://surfer.nmr.mgh.harvard.edu/). Any sphere sections that overlapped with the cortex (no lower limit restrictions) were mapped to the corresponding parcel, and we tallied how many studies had positive indicator values in each parcel. Two studies from the subjective valuation corpus were removed because they did not contain cortical voxels, leaving a final number of 198 studies.

With these lists, we tested for parcels that were significantly more strongly associated with one domain than the other. We permuted the label of every study (i.e. DMN or valuation) 5000 times while preseving the total number of studies observed in each domain, and on each iteration stored the maximum resulting statistic (across all parcels) from per-parel chi-squared tests comparing the proportion of studies with activation in that parcel in the two domains. This gave us a null distribution of 5000 maximum chi-squared values. The 95th percentile of this distribution served as an FWE-corrected significance threshold to evaluate unpermuted chi-squared values.

@@

Our fMRI analyses used resting-state fMRI data from the Human Connectome Project [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state runs (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) each comprised 1200 time points(14 min 33 s) for a total of 4800 time points. Two runs were acquired on each day, with the phase encoding direction set to left-right for one run and right-left for the other. Only subjects with both left-right and right-left phase encoding were included (i.e. subjects with four rsfMRI sessions). In addition, only datasets with either low motion levels (under 1.5mm) or under 0.5 mm mean framewise displacement (FD) were used. See [@VanEssen2012] for more details about the data acquisition protocol.

Data initially underwent the HCP minimal preprocessing pipeline [@Glasser2013], which included gradient nonlinearity correction, motion correction, EPI distortion correction, temporal denoising based on independent components (FIX, [@Griffanti2014; @Salimi-Khorshidi2014]), high-pass filtering (0.0005 Hz threshold), and MNI152-based normalization. Data were further preprocessed using an in-house pipeline described previously [@Tobyne2017]. Steps included linear interpolation across timepoints with over 0.5 mm of framewise displacement, band-pass filtering (allowed frequencies ranged from 0.009 and 0.08 Hz), temporal denoising and high-motion censoring by deletion, as well as mean greyordinate signal regression [@Burgess2016]. Volumes that displayed an FD of over 0.5 mm were considered as spikes and removed from the subject's dataset, and subjects whose spike counts were above 5% of their total data were excluded from analyses. Finally, data acquired on the same day (i.e. left-right and right-left phase encoding session) were temporally demeaned. Each subject's brain was comprised of 32k standard HCP greyordinates per hemisphere (combined in a CIFTI file). We retained only the cortical surfaces, which resulted in 59,412 total surface vertices per subject.

### *Network Definition*

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. We report the proportion of retained edges (i.e. network density) in the Results section. As part of the evaluation step, we produced these network weight matrices (and the community detection outlined below) at four levels: 1) for each run separately (1200 TRs each); 2) for the concatenated time series for the two runs on the same day (2400 TRs total); 3) for each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

We performed a coordinate-based meta-analysis to identify unique standard regions across DMN and SV literatures. Volumetric coordinates from a set of DMN and SV studies were projected onto a cortical surface, and mapped to a multimodal cortical parcellation [@Glasser2016] to produce a list of brain areas reported per study. The 40 areas considered were limited to the medial portion of the default mode and limbic networks, as defined by the Yeo et al.'s [-@Yeo2011] 17 network parcellation. Unique areas were identified by first permuting the affiliation of each study (i.e. DMN or SV) to create a distribution of chi-squared statistics based on the comparison of the number of reports per area between literatures (see Methods for details). This putative null distribution was used to identify regions more often reported in either literature.
